---
title: "Assignment 4"
author: "Angie Bouche, Andrea Cheung, Tara Jagadeesh"
date: "November 19, 2018"
output: html_document
---

```{r setup,echo = FALSE, message = FALSE}

#load in packages and data as csvs
library(tidyverse)
library(knitr)
library(kableExtra)
library(effsize)
library(reshape2)

lobster_traps <- read.csv("lobster_traps.csv")
lobster_size <- read.csv("lobster_size_abundance.csv")

lobster_expanded <- expand.dft(lobster_size, freq="COUNT")
traps_expanded <- expand.dft(lobster_traps, freq="TRAPS")


```



```{r}
#Make a summary data table for lobster abundance 

lobster_count <- lobster_expanded %>% 
  filter(SIZE != -99999) %>% #remove negative value
  select (YEAR, SITE) %>% #select relevant years
   group_by(SITE, YEAR) %>% 
  summarize(
    Sum = length(SITE))
 
```

```{r}
#Make a summary data table for fishing pressure
trap_count <- traps_expanded %>% 
 select (YEAR, SITE) %>% #select relevant columns
  group_by (SITE, YEAR) %>%
  filter(SITE != "ABUR") %>%
  filter(SITE != "AHND to AQUE") %>% 
  filter(SITE != "AHND") %>%  
  filter(SITE != "GOLB") %>%  
  summarize(
    Sum = length(SITE))

```



```{r}
# Make a line graph of annual lobster abundance at each site

counts_line <- ggplot (lobster_count, aes(x = YEAR, y = Sum)) +
  geom_line(aes(fill = SITE))+
  scale_x_continuous(limits = c(2012,2017)) +
  theme(plot.title = element_text(hjust = 0.5))+
  labs( y = "Lobster Abundance", x = "Year")
counts_line

```


```{r}
#Make a line graph of annual lobster fishing pressure at each site

trap_line <- ggplot (trap_count, aes(x = YEAR, y = Sum)) +
  geom_line(aes(fill = SITE))+
  scale_x_continuous(limits = c(2012,2017)) +
  theme(plot.title = element_text(hjust = 0.5))+
  labs( y = "Number of Traps", x = "Year")
trap_line

```



## Number 2
Create data frame where each lobster has its own row, case format using vcdExtra
```{r}

library(vcdExtra)

lobster_expanded <- expand.dft(lobster_size, freq="COUNT")

lobster_expanded2017 <- lobster_expanded %>% 
  filter(YEAR == 2017) 


```

Exploratory Histogram 
```{r}
size2017_hist <- ggplot(lobster_expanded, aes(x=SIZE)) +
  geom_histogram()+
  facet_wrap(~SITE, scale = "free")

lobstersize_hist <- ggplot(lobster_expanded2017, aes(x=SIZE)) +
  geom_histogram()


size2017_hist

#large sample size, normally distributed
```

Exploratory qqplot

```{r}

size2017_qq <- ggplot(lobster_expanded, aes(sample=SIZE)) +
  geom_qq()+
  facet_wrap(~SITE, scale = "free")

lobstersize_qq <- ggplot(lobster_expanded2017, aes(sample=SIZE)) +
  geom_qq()

size2017_qq

#large sample size and normally distributed according to qqplot

```

Variance test (Levene's)
```{r}
# install car package to compute levene's test
library(car)

# Levene's test with one independent variable
leveneTest(SIZE ~ SITE, data = lobster_expanded2017)

# low P value, according to the pvalue, variance is not equal, but its less than 4x. 

```

One way ANOVA - is there a difference between two or more means?
```{r}
lobstersize_aov <- aov(SIZE ~ SITE, data = lobster_expanded2017)
lobstersize_aov

summary(lobstersize_aov)
```

<<<<<<< HEAD

## Number 2
Create data frame where each lobster has its own row, case format using vcdExtra
```{r}

library(vcdExtra)

lobster_expanded <- expand.dft(lobster_size, freq="COUNT")

lobster_expanded2017 <- lobster_expanded %>% 
  filter(YEAR == 2017) 


```

Exploratory Histogram 
```{r}
lobstersize_hist <- ggplot(lobster_expanded2017, aes(x=SIZE)) +
  geom_histogram()

lobstersize_hist

#large sample size, normally distributed
```

Exploratory qqplot

```{r}
lobstersize_qq <- ggplot(lobster_expanded2017, aes(sample=SIZE)) +
  geom_qq()

lobstersize_qq

#large sample size and normally distributed according to qqplot

```

Variance test (Levene's)
```{r}
# install car package to compute levene's test
library(car)

# Levene's test with one independent variable
leveneTest(SIZE ~ SITE, data = lobster_expanded2017)

# low P value, according to the pvalue, variance is not equal, but its less than 4x. 

```

One way ANOVA - is there a difference between two or more means?
```{r}
lobstersize_aov <- aov(SIZE ~ SITE, data = lobster_expanded2017)
lobstersize_aov

summary(lobstersize_aov)
```


Post-hoc Tukey's test
```{r}
tukey <- TukeyHSD(lobstersize_aov)

tukey
```

Graph data
```{r}
library(ggplot2)

lobster_size_2017 <- lobster_expanded2017 %>% 
  filter(YEAR == 2017) %>% 
  group_by(SITE) %>% 
  summarize(meansize = mean(SIZE),
            sd=round(sd(SIZE), digits = 2),
            sample_size = length(SITE))

lobstersize2017_figure <-ggplot(lobster_size_2017, aes(x=SITE, y= meansize, fill = SITE)) + 
  geom_col(colour = "black", width = 0.5, stat = "identity") +
  labs(x = "Location", y = "Carapace Length (mm)", title = "Mean carapace length (mm) of \nlobsters by location in 2017") +
  geom_errorbar(aes(ymin=meansize-sd, ymax=meansize+sd), width=.2)+
  theme_classic()+ 
  scale_y_continuous(expand = c(0,0), limits = c(0,125))+
  theme(legend.position="none", aspect.ratio = 1.2/1)
  
# Mean carapace length (mm) of lobsters at five locations: Arroyo Quemado (AQUE), Carpinteria (CARP), Isla Vista (IVEE), Mohawk Reef (MOHK), Naples Reef (NAPL). Error bars indicate ± 1 standard deviation (n=67, n=705, n=606, n=178, n=112, respectively). Mean carapace length of lobsters only differed significantly between NAPL and IVEE, and NAPL and CARP. (one-way ANOVA, F(4,1663) = 3.42, p = 0.0085) with post-hoc Tukey’s HSD (α = 0.05). 


lobstersize2017_figure
```

Boxplot
```{r}

lobstersize2017_box <-ggplot(lobster_expanded2017, aes(x=SITE, y= SIZE, fill = SITE)) + 
  geom_boxplot() +
  labs(x = "Location", y = "Carapace Length (mm)", title = "Mean carapace length (mm) of lobsters \nby location in 2017", scale = "free") +
  theme(legend.position="none", aspect.ratio = 1.2/1) 


lobstersize2017_box
```


## Number 3
Histogram + qqplots for 2012, and F tests 
```{r}
size2012_hist <- lobster_expanded %>% 
  filter(YEAR == "2012") %>% 
  ggplot(aes(x = SIZE)) +
  geom_histogram()+
  facet_wrap(~SITE, scale = "free")
size2012_hist

size2012_qq <- lobster_expanded %>% 
  filter(YEAR == "2012") %>% 
  ggplot(aes(sample=SIZE)) +
  geom_qq()+
  facet_wrap(~SITE, scale = "free")
size2012_qq

#F tests
IVEE_ftest <- size_simple %>% 
  filter(SITE == "IVEE") %>% 
  var.test(SIZE ~ YEAR, data = .)
IVEE_ftest #Equal

NAPL_ftest <- size_simple %>% 
  filter(SITE == "NAPL") %>% 
  var.test(SIZE ~ YEAR, data = .)
NAPL_ftest #Equal

AQUE_ftest <- size_simple %>% 
  filter(SITE == "AQUE") %>% 
  var.test(SIZE ~ YEAR, data = .)
AQUE_ftest #Equal

CARP_ftest <- size_simple %>% 
  filter(SITE == "CARP") %>% 
  var.test(SIZE ~ YEAR, data = .)
CARP_ftest #Equal

MOHK_ftest <- size_simple %>% 
  filter(SITE == "MOHK") %>% 
  var.test(SIZE ~ YEAR, data = .)
MOHK_ftest #Equal


```



```{r}
#Question: is there a difference in mean lobster size between 2012 and 2017?


size_simple <- lobster_expanded %>% 
  filter(YEAR == "2012" | YEAR == "2017") %>% 
  select("YEAR", "SITE", "SIZE")
  
## T tests and effect size for each site

#IVEE 
ivee_ttest <- size_simple %>% 
  filter(SITE == "IVEE") %>% 
  t.test(SIZE ~ YEAR, var.equal = TRUE, data = .)
ivee_ttest

ivee_effectsize <- size_simple %>% 
  filter(SITE == "IVEE") %>% 
  cohen.d(SIZE ~ YEAR, data = .)
ivee_effectsize #0.3775177 (small)

#NAPL
napl_ttest <- size_simple %>% 
  filter(SITE == "NAPL") %>% 
  t.test(SIZE ~ YEAR, var.equal = TRUE, data = .)
napl_ttest

napl_effectsize <- size_simple %>% 
  filter(SITE == "NAPL") %>% 
  cohen.d(SIZE ~ YEAR, data = .)
napl_effectsize #0.2834216 (small)

#AQUE 
aque_ttest <- size_simple %>% 
  filter(SITE == "AQUE") %>% 
  t.test(SIZE ~ YEAR, var.equal = TRUE, data = .)
aque_ttest

aque_effectsize <- size_simple %>% 
  filter(SITE == "AQUE") %>% 
  cohen.d(SIZE ~ YEAR, data = .)
aque_effectsize #0.2563169 (small)

#CARP 
carp_ttest <- size_simple %>% 
  filter(SITE == "CARP") %>% 
  t.test(SIZE ~ YEAR, var.equal = TRUE, data = .)
carp_ttest

carp_effectsize <- size_simple %>% 
  filter(SITE == "CARP") %>% 
  cohen.d(SIZE ~ YEAR, data = .)
carp_effectsize #0.1594364 (negligible)

#MOHK
mohk_ttest <- size_simple %>% 
  filter(SITE == "MOHK") %>% 
  t.test(SIZE ~ YEAR, var.equal = TRUE, data = .)
mohk_ttest

mohk_effectsize <- size_simple %>% 
  filter(SITE == "MOHK") %>% 
  cohen.d(SIZE ~ YEAR, data = .)
mohk_effectsize #0.5408116 (medium)

```


Graph of results
```{r}
#Summary table of mean lobster size for MPA and NONMPA sites in 2012 and 2017
size_means <- size_simple %>% 
  group_by(SITE, YEAR) %>% 
  summarize(
    Mean = mean(SIZE)
  )
size_means



size_graph <- size_means %>% 
  ggplot(aes(factor(YEAR), Mean, fill = SITE))+
  geom_bar(stat="identity", position = "dodge") 
size_graph
```



##Number 4 
Chi Square

```{r}



lobster_chitable <- lobster_expanded2017 %>% 
  filter(YEAR == 2017) %>% 
  mutate(
    Size = case_when(
      SIZE >= 82.6 ~ "greater than",
      SIZE <= 82.6 ~ "less than")) %>% 
  select(SITE, Size) %>% 
  count(Size, SITE) %>% 
  spread(SITE, n) %>% 
  select(-Size)


rownames(lobster_chitable) <- c("greater than", "less than")
lobster_chitable


lobster_x2 <- chisq.test(lobster_chitable) #run x2 test
lobster_x2


```